An Environment for Sustainable Research Software in Germany and Beyond: Current State, Open Challenges, and Call for Action

### 2020-04-05 001/manuscript.tex
Add glossary and move Suppl Mat to new page
Update manuscript.tex
Minor corrections discovered during preparation of journal submission

% Custom packages
\usepackage[nonumberlist]{glossaries} % TODO Check if we can put glossary items in the abbreviations section
%\usepackage[inline]{enumitem}
% Set graphics path
\graphicspath{{figures/}}
% EO Custom packages

@@

%%% Use the \authfn to add symbols for additional footnotes, if any. 1 is reserved for correspondence emails; then continuing with 2 etc for contributions.
\author[1,2,\authfn{2}]{Hartwig Anzt} % 0000-0003-2177-952X
\author[1,\authfn{2}]{Felix Bach} % 0000-0002-5035-7978
\author[3,4,5,\authfn{2}]{Stephan Druskat} % 0000-0003-4925-7248
\author[6,7,8,\authfn{2}]{Frank L\"offler} % 0000-0001-6643-6323
\author[9,\authfn{1},\authfn{2}]{Axel Loewe} % 0000-0002-2487-4744
\author[10,11\authfn{2}]{Bernhard Y.\ Renard} % 0000-0003-4589-9809
\author[12,13,\authfn{1}\authfn{2}]{Gunnar Seemann} % 0000-0001-7111-7992
\author[14,\authfn{2}]{Alexander Struck} % 0000-0002-1173-9228
\author[15]{Elke Achhammer}
\author[16]{Piush Aggarwal}
\author[17]{Franziska Appel}
\author[18]{Michael Bader}
\author[19]{Lutz Brusch}
\author[20]{Christian Busse}
\author[21]{Gerasimos Chourdakis}
\author[22]{Peter Ebert}
\author[23]{Bernd Flemisch}
\author[24]{Sven Friedl}
\author[25]{Bernadette Fritzsch}
\author[26]{Maximilian D.\ Funk}
\author[3]{Volker Gast}
\author[27]{Florian Goth}
\author[28]{Jean-No\"{e}l Grad}
\author[29]{Sibylle Hermann}
\author[30]{Florian Hohmann}
\author[31]{Stephan Janosch}
\author[32]{Dominik Kutra}
\author[33]{Jan Linxweiler}
\author[10,34]{Thilo Muth}
\author[35]{Wolfgang Peters-Kottig}
\author[36]{Fabian Rack}
\author[37]{Fabian H.C.\ Raters}
\author[38]{Stephan Rave}
\author[39]{Guido Reina}
\author[40]{Malte Reißig}
\author[41,42]{Timo Ropinski}
\author[43]{Joerg Schaarschmidt}
\author[44,45,46]{Heidi Seibold}
\author[47]{Jan P.\ Thiele}
\author[48]{Benjamin Uekerman}
\author[49]{Stefan Unger}
\author[28]{Rudolf Weeber}
\author[50]{Piotr Wojciech Dabrowski}

@@

\affil[22]{Center for Bioinformatics Saar, Saarland Informatics Campus, Germany}
\affil[23]{Institute for Modelling Hydraulic and Environmental Systems, University of Stuttgart, Germany}
\affil[24]{Berlin Institute of Health, Germany}
\affil[25]{Scientific Computing, Alfred Wegener Institute, Helmholtz Center for Polar and Marine Research Bremerhaven, Germany}
\affil[26]{Max-Planck-Gesellschaft e.V.}
\affil[27]{Institut für Theoretische Physik und Astrophysik, Universität Würzburg, Germany}
\affil[28]{Institut für Computerphysik, University of Stuttgart, Germany}
\affil[29]{University Library, University of Stuttgart, Germany}
\affil[30]{Zentrum für Medien-, Kommunikations- und Informationsforschung (ZeMKI), Universität Bremen, Germany}
\affil[31]{Max Planck Institue of Molecular Cell Biology and Genetics, Dresden, Germany}
\affil[32]{European Molecular Biology Laboratory, Heidelberg, Germany}
\affil[33]{Center for Mechanics, Uncertainty and Simulation in Engineering, Technische Universität Braunschweig, Germany}
\affil[34]{eScience Division, Federal Institute for Materials Research and Testing, Berlin, Germany}
\affil[35]{Konrad-Zuse-Zentrum für Informationstechnik Berlin (ZIB), Germany}
\affil[36]{FIZ Karlsruhe - Leibniz Institute for Information Infrastructure, Germany}
\affil[37]{Department of Economics, University of Goettingen, Germany}
\affil[38]{Applied Mathematics, University of Münster, Germany}
\affil[39]{Visualization Research Center, University of Stuttgart, Germany}
\affil[40]{Institute for Advanced Sustainability Studies e.V.}
\affil[41]{Institute of Media Informatics, Ulm University, Germany}
\affil[42]{Department of Science and Technology, Linköping University, Sweden}
\affil[43]{Institute of Nanotechnology, Karlsruhe Institute of Technology (KIT),  Germany}
\affil[44]{LMU Munich, Germany}
\affil[45]{Bielefeld University, Germany}
\affil[46]{Helmholtz Zentrum Munich, Germany}
\affil[47]{Institute of Applied Mathematics, Leibniz University Hannover, Germany}
\affil[48]{Energy Technology, Eindhoven University of Technology, The Netherlands}
\affil[49]{Julius Kühn-Institut (JKI), Federal Research Centre for Cultivated Plants, Quedlinburg, Germany}
\affil[50]{School of Computing, Communication and Business, Hochschule für Technik und Wirtschaft Berlin, Germany}

@@

% Glossaries
\makeglossaries

@@

\subsection{Ideas for Solutions}
In order to meet the legal challenges mentioned, it is absolutely necessary for the software developer (team) to document the rights chain comprehensively during the software development (see ,e.g., supplementary material). Contributions of individual persons must be traceable and their (labor law) status must be named. At best, contracts with rules on the transfer of rights of use should be concluded before work begins. Declarations of assignment of rights can be made for existing works. License conditions for external contributions must be evaluated with regard to further rights of use and possible sub-licensing. Contracts and funding conditions must be conscientiously documented and analyzed with regard to rules on rights of use. In case that different parts of the software are based on different conditions and rights of third parties, individual modules of the new software could be published under different licenses and merged accordingly.

@@

\printglossaries

@@

\subsection{Author's Contributions}
We are a group of software-providing researchers, RSEs, and infrastructural as well as legal supporters. Initially, a group of representatives of funded projects of the first DFG sustainability call~\cite{DFGfunded} met during the first German RSE conference (deRSE19)~\cite{deRSEconf} in June 2019 in a grass-roots workshop on sustainable research software addressing the software-based research community. During this workshop, we realized that a lot of valuable experience and good ideas are present in the group, and we decided to start working on this paper together with other interested practitioners. We followed the generous invitation of the DFG for the above-mentioned two-day meeting at the Robert Koch Institute in Berlin in November 2019 to sharpen the focus of this paper.\\
The individual contributions of authors to this work are detailed below, following the \href{https://casrai.org/credit/}{CASRAI CRediT} (Contributor Roles Taxonomy):
% TODO Authors, please add yourself to this list under Writing - original draft
\begin{itemize}
    \item \textbf{Conceptualization:}  Anzt, Bach, Druskat, Loewe, Löffler,
    Renard, Seemann, Struck
    \item \textbf{Funding acquisition:} Loewe, Seemann
    \item \textbf{Investigation:} Bach, Druskat, Loewe, Löffler, Renard,
    Seemann
    \item \textbf{Project administration:} Loewe, Seemann
    \item \textbf{Visualization:} Unger, Friedl, Löffler
    \item \textbf{Writing original draft:} Achhammer, Aggarwal, Anzt, Appel, Bach, Bader, Brusch, Druskat, Ebert, Flemisch, Friedl, Funk, Grad, Goth, Herrmann, Hohmann, Kutra, Linxweiler, Loewe,
    Löffler, Muth, Peters-Kottig, Rack, Raters, Rave, Reina, Renard, Ropinksi, Schaarschmidt, Seemann, Struck, Thiele, Uekermann, Unger, Weeber
    \item \textbf{Writing review \& editing:} Anzt, Appel, Bach, Bader, Brusch, 
    Busse, Chourdakis, Druskat, Friedl, Fritzsch, Funk, Gast, Herrmann, Janosch, Loewe, Löffler, Rack, Reina,
    Renard, Seemann, Seibold, Struck, Thiele, Uekermann, Wojciech Dabrowski
\end{itemize}

@@

@@

\begin{figure}[tb]
  \centering
  \includegraphics[scale=0.5]{tree_IP_policy}
  \caption{\textbf{Policy}. This tree recommends to check closely any policies implemented in the software developers organization.}
  \label{fig:tree0}
\end{figure}

\begin{figure}[tb]
  \centering
  \includegraphics[scale=0.5]{figures/tree_sustainable_software}
  \caption{\textbf{Contributors}. This tree helps to find out whether the academic institution where the software development is located is the owner of the intellectual property (copyright).}
  \label{fig:tree1}
\end{figure}

\begin{figure}[tb]
  \centering
  \includegraphics[scale=0.5]{tree_new_SW_development}
  \caption{\textbf{Code history}. The code history tree points out tasks for projects that incorporate existing code.}
  \label{fig:tree2}
\end{figure}

\begin{figure}[tb]
  \centering
  \includegraphics[scale=0.5]{tree_licensing}
  \caption{\textbf{Licensing}. Depending on the distribution model, open access (OA) or open source software can be selected.}
  \label{fig:tree3}
\end{figure}

>>>>

% Custom packages
\usepackage[nonumberlist]{glossaries} % TODO Check if we can put glossary items in the abbreviations section
% \usepackage{flushend}
%\usepackage[inline]{enumitem}
% Set graphics path
\graphicspath{{figures/}}
% EO Custom packages

@@

%%% Use the \authfn to add symbols for additional footnotes, if any. 1 is reserved for correspondence emails; then continuing with 2 etc for contributions.
\author[1,2,\authfn{2}]{Hartwig Anzt} % 0000-0003-2177-952X
\author[1,\authfn{2}]{Felix Bach} % 0000-0002-5035-7978
\author[3,4,5,\authfn{2}]{Stephan Druskat} % 0000-0003-4925-7248
\author[6,7,8,\authfn{2}]{Frank L\"offler} % 0000-0001-6643-6323
\author[9,\authfn{1},\authfn{2}]{Axel Loewe} % 0000-0002-2487-4744
\author[10,11\authfn{2}]{Bernhard Y.\ Renard} % 0000-0003-4589-9809
\author[12,13,\authfn{1}\authfn{2}]{Gunnar Seemann} % 0000-0001-7111-7992
\author[14,\authfn{2}]{Alexander Struck} % 0000-0002-1173-9228
\author[15]{Elke Achhammer}
\author[16]{Piush Aggarwal}
\author[17]{Franziska Appel}
\author[18]{Michael Bader}
\author[19]{Lutz Brusch}
\author[20]{Christian Busse}
\author[21]{Gerasimos Chourdakis}
\author[22]{Piotr W. Dabrowski}
\author[23]{Peter Ebert}
\author[24]{Bernd Flemisch}
\author[25]{Sven Friedl}
\author[26]{Bernadette Fritzsch}
\author[27]{Maximilian D.\ Funk}
\author[3]{Volker Gast}
\author[28]{Florian Goth}
\author[29]{Jean-No\"{e}l Grad}
\author[30]{Sibylle Hermann}
\author[31]{Florian Hohmann}
\author[32]{Stephan Janosch}
\author[33]{Dominik Kutra}
\author[34]{Jan Linxweiler}
\author[10,35]{Thilo Muth}
\author[36]{Wolfgang Peters-Kottig}
\author[37]{Fabian Rack}
\author[38]{Fabian H.C.\ Raters}
\author[39]{Stephan Rave}
\author[40]{Guido Reina}
\author[41]{Malte Reißig}
\author[42,43]{Timo Ropinski}
\author[44]{Joerg Schaarschmidt}
\author[45,46,47]{Heidi Seibold}
\author[48]{Jan P.\ Thiele}
\author[49]{Benjamin Uekerman}
\author[50]{Stefan Unger}
\author[29]{Rudolf Weeber}

@@

\affil[22]{School of Computing, Communication and Business, Hochschule für Technik und Wirtschaft Berlin, Germany}
\affil[23]{Center for Bioinformatics Saar, Saarland Informatics Campus, Germany}
\affil[24]{Institute for Modelling Hydraulic and Environmental Systems, University of Stuttgart, Germany}
\affil[25]{Berlin Institute of Health, Germany}
\affil[26]{Scientific Computing, Alfred Wegener Institute, Helmholtz Center for Polar and Marine Research Bremerhaven, Germany}
\affil[27]{Max-Planck-Gesellschaft e.V.}
\affil[28]{Institut für Theoretische Physik und Astrophysik, Universität Würzburg, Germany}
\affil[29]{Institut für Computerphysik, University of Stuttgart, Germany}
\affil[30]{University Library, University of Stuttgart, Germany}
\affil[31]{Zentrum für Medien-, Kommunikations- und Informationsforschung (ZeMKI), Universität Bremen, Germany}
\affil[32]{Max Planck Institute of Molecular Cell Biology and Genetics, Dresden, Germany}
\affil[33]{European Molecular Biology Laboratory, Heidelberg, Germany}
\affil[34]{Center for Mechanics, Uncertainty and Simulation in Engineering, Technische Universität Braunschweig, Germany}
\affil[35]{eScience Division, Federal Institute for Materials Research and Testing, Berlin, Germany}
\affil[36]{Konrad-Zuse-Zentrum für Informationstechnik Berlin (ZIB), Germany}
\affil[37]{FIZ Karlsruhe - Leibniz Institute for Information Infrastructure, Germany}
\affil[38]{Department of Economics, University of Goettingen, Germany}
\affil[39]{Applied Mathematics, University of Münster, Germany}
\affil[40]{Visualization Research Center, University of Stuttgart, Germany}
\affil[41]{Institute for Advanced Sustainability Studies e.V.}
\affil[42]{Institute of Media Informatics, Ulm University, Germany}
\affil[43]{Department of Science and Technology, Linköping University, Sweden}
\affil[44]{Institute of Nanotechnology, Karlsruhe Institute of Technology (KIT),  Germany}
\affil[45]{LMU Munich, Germany}
\affil[46]{Bielefeld University, Germany}
\affil[47]{Helmholtz Zentrum Munich, Germany}
\affil[48]{Institute of Applied Mathematics, Leibniz University Hannover, Germany}
\affil[49]{Energy Technology, Eindhoven University of Technology, The Netherlands}
\affil[50]{Julius Kühn-Institut (JKI), Federal Research Centre for Cultivated Plants, Quedlinburg, Germany}

@@

% Glossaries
% \makeglossaries
\makenoidxglossaries

@@

\subsection{Ideas for Solutions}
In order to meet the legal challenges mentioned, it is absolutely necessary for the software developer (team) to document the rights chain comprehensively during the software development (see, e.g., supplementary material). Contributions of individual persons must be traceable and their (labor law) status must be named. At best, contracts with rules on the transfer of rights of use should be concluded before work begins. Declarations of assignment of rights can be made for existing works. License conditions for external contributions must be evaluated with regard to further rights of use and possible sub-licensing. Contracts and funding conditions must be conscientiously documented and analyzed with regard to rules on rights of use. In case that different parts of the software are based on different conditions and rights of third parties, individual modules of the new software could be published under different licenses and merged accordingly.

@@

% \printglossaries
\printnoidxglossaries

@@

\subsection{Author's Contributions}
We are a group of software-providing researchers, RSEs, and infrastructural as well as legal supporters. Initially, a group of representatives of funded projects of the first DFG sustainability call~\cite{DFGfunded} met during the first German RSE conference (deRSE19)~\cite{deRSEconf} in June 2019 in a grass-roots workshop on sustainable research software addressing the software-based research community. During this workshop, we realized that a lot of valuable experience and good ideas are present in the group, and we decided to start working on this paper together with other interested practitioners. We followed the generous invitation of the DFG for the above-mentioned two-day meeting at the Robert Koch Institute in Berlin in November 2019 to sharpen the focus of this paper.\\
The individual contributions of authors to this work are detailed below, following the \href{https://casrai.org/credit/}{CASRAI CRediT} (Contributor Roles Taxonomy):
% TODO Authors, please add yourself to this list under Writing - original draft
\begin{itemize}
    \item \textbf{Conceptualization:}  Anzt, Bach, Druskat, Loewe, Löffler,
    Renard, Seemann, Struck
    \item \textbf{Funding acquisition:} Loewe, Seemann
    \item \textbf{Investigation:} Bach, Druskat, Loewe, Löffler, Renard,
    Seemann
    \item \textbf{Project administration:} Loewe, Seemann
    \item \textbf{Visualization:} Unger, Friedl, Löffler
    \item \textbf{Writing original draft:} Achhammer, Aggarwal, Anzt, Appel, Bach, Bader, Brusch, Druskat, Ebert, Flemisch, Friedl, Funk, Grad, Goth, Herrmann, Hohmann, Kutra, Linxweiler, Loewe,
    Löffler, Muth, Peters-Kottig, Rack, Raters, Rave, Reina, Renard, Ropinksi, Schaarschmidt, Seemann, Struck, Thiele, Uekermann, Unger, Weeber
    \item \textbf{Writing review \& editing:} Anzt, Appel, Bach, Bader, Brusch, 
    Busse, Chourdakis, Dabrowski, Druskat, Friedl, Fritzsch, Funk, Gast, Herrmann, Janosch, Loewe, Löffler, Rack, Reina, Reißig,
    Renard, Seemann, Seibold, Struck, Thiele, Uekermann, 
\end{itemize}

@@

\clearpage

@@

\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.5]{tree_IP_policy}
  \caption{\textbf{Policy}. This tree recommends to check closely any policies implemented in the software developers organization.}
  \label{fig:tree0}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.5]{figures/tree_sustainable_software}
  \caption{\textbf{Contributors}. This tree helps to find out whether the academic institution where the software development is located is the owner of the intellectual property (copyright).}
  \label{fig:tree1}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.5]{tree_new_SW_development}
  \caption{\textbf{Code history}. The code history tree points out tasks for projects that incorporate existing code.}
  \label{fig:tree2}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.5]{tree_licensing}
  \caption{\textbf{Licensing}. Depending on the distribution model, open access (OA) or open source software can be selected.}
  \label{fig:tree3}
\end{figure}
